Dynamics of Laterally Propagating Flames in X-ray Bursts. I. Burning Front Structure

### 2020-02-25 xrb/submit/paper.tex
Change coordinate in HSE equation for lower boundary
latest changes
Fixed typo, added a few comments to diff

@@

Since the neutron star rotates, we work in a corotating frame, taking
the angular velocity $\Omegab$ to be constant.  Further, for the two-dimensional
simulations presented here, we work in axisymmetric coordinates,
but we advect a third component of velocity, coming out of the
simulation plane, that participates in the Coriolis force (sometimes
described as a 2.5D simulation).  We will take $r$ to be the cylindrical
radial coordinate with corresponding velocity $u$, $z$ to be the
cylindrical vertical coordinate with corresponding velocity $v$, and the
$\theta$ to be the cylindrical azimuthal coordinate with corresponding velocity
$w$. As $r$ and $z$ define the simulation plane, we adopt the nonstandard ordering
($r$, $z$, $\theta$), for which a righthanded coordinate system has positive $w$
pointing out of the page.  We will take $\Omegab = \Omega_0 \hat{\bf z}$ for the
angular rotation rate, and $\gb = -g \hat{\bf z}$ for the gravitational acceleration,
with $g$ constant. With these choices, the Coriolis force is:
\begin{equation}
-2\rho \Omegab \times \Ub =
   -2\rho \left ( \Omega_0 w \hat{\bf r} - \Omega_0 u \hat{\boldsymbol{\theta}} \right )
\end{equation}
We will neglect the centrifugal force---with our plane-parallel
geometry, this will act only in the lateral direction, and is not
expected to greatly affect the dynamics.  Carrying the Coriolis force
allows us to capture the geostrophic balance that sets up via lateral
hydrostatic equilibrium \citep{spitkovsky2002}.

@@

Thermal diffusion is modeled explicitly, using a predictor-corrector
scheme to achieve second-order accuracy.  A verification test of the
diffusion scheme is shown in Appendix~\ref{app:diffusion}.  The
explicit thermal diffusion requires a timestep limiter of the form:
\begin{equation}
\Delta t_\mathrm{diff} \le \frac{1}{2} \frac{\Delta x^2}{\mathcal{D}}
\end{equation}
where $\mathcal{D} = \kth/(\rho c_v)$ is the thermal diffusivity.  The
diffusivity increases rapidly at the top of the atmosphere, causing
these low density regions to determine the overall timestep for the
simulations.  Therefore, we disable thermal conduction at low
densities where it is not expected to be important.

We use hydrostatic boundary conditions on the lower boundary, using
a discretized hydrostatic equilibrium equation of the form:
\begin{equation}
\label{eq:hse}
    p_i = p_{i-1} + \frac{1}{2} \Delta y (\rho_i + \rho_{i-1}) \gb \cdot \hat{\bf y}
\end{equation}
and holding the temperature constant in the ghost cells.  This is solved
together with the equation of state.  The velocity is reflected at this boundary.
This procedure follows the form described in \citet{ppm-hse}.  The
left boundary is reflecting and the right boundary is a zero-gradient
outflow.  The top boundary sets the state to simply the conditions
in our outer buffer region of the initial model (see below), with the
normal velocity set to the larger of zero or the velocity at the top
of the domain (this prevents incoming velocities at the top) and the
transverse velocities set to zero.

@@

%======================================================================
% Initial Model
%======================================================================
\section{Initial Model}\label{Sec:inital_model}

@@

\added{
\citet{cavecchi:2013} ran simulations at rotation rates of $450$ Hz exploring a range of opacities.
Accounting for the boosting, we would expect to see our 10/10 flame propagate at $\sim 2.25$ times
the speed of these $450$ Hz runs, but instead the speeds range from $4.5$ to $7$ times those reported in
\citet{cavecchi:2013}. The additional boost may come from cyclic fluid motions that would have been
disallowed in the $450$ Hz runs by the enforcement of vertical hydrostatic equilibrium (we already saw
that hydrodynamic effects can accelerate the burning front in our comparison with the 1D flames). The
inclusion of reaction chains beyond may also increase energy generation and consequently the
propagation speed.}

>>>>

\explain{We've updated the equations in this section to use cylindrical
coordinates.}

@@

Since the neutron star rotates, we work in a corotating frame, taking
the angular velocity $\Omegab$ to be constant.  Further, for the two-dimensional
simulations presented here, we work in axisymmetric coordinates,
but we advect a third component of velocity, coming out of the
simulation plane, that participates in the Coriolis force (sometimes
described as a 2.5D simulation).  We will take $r$ to be the cylindrical
radial coordinate with corresponding velocity $u$, $z$ to be the
cylindrical vertical coordinate with corresponding velocity $v$, and the
$\theta$ to be the cylindrical azimuthal coordinate with corresponding velocity
$w$. \added{As $r$ and $z$ define the simulation plane, we adopt the nonstandard ordering
($r$, $z$, $\theta$)}, for which a righthanded coordinate system has positive $w$
pointing out of the page.  We will take $\Omegab = \Omega_0 \hat{\bf z}$ for the
angular rotation rate, and $\gb = -g \hat{\bf z}$ for the gravitational acceleration,
with $g$ constant. With these choices, the Coriolis force is:
\begin{equation}
-2\rho \Omegab \times \Ub =
   -2\rho \left ( \Omega_0 w \hat{\bf r} - \Omega_0 u \hat{\boldsymbol{\theta}} \right )
\end{equation}
We will neglect the centrifugal force---with our plane-parallel
geometry, this will act only in the lateral direction, and is not
expected to greatly affect the dynamics.  Carrying the Coriolis force
allows us to capture the geostrophic balance that sets up via lateral
hydrostatic equilibrium \citep{spitkovsky2002}.

@@

Thermal diffusion is modeled explicitly, using a predictor-corrector
scheme to achieve second-order accuracy.  A verification test of the
diffusion scheme is shown in Appendix~\ref{app:diffusion}.  The
explicit thermal diffusion requires a timestep limiter of the form:
\begin{equation}
\Delta t_\mathrm{diff} \le \frac{1}{2} \frac{h^2}{\mathcal{D}}
\end{equation}
where $\mathcal{D} = \kth/(\rho c_v)$ is the thermal diffusivity and $h$ is the width of a cell (either $\Delta r$ or $\Delta z$).  The
diffusivity increases rapidly at the top of the atmosphere, causing
these low density regions to determine the overall timestep for the
simulations.  Therefore, we disable thermal conduction at low
densities where it is not expected to be important.

We use hydrostatic boundary conditions on the lower boundary, using
a discretized hydrostatic equilibrium equation of the form:
\begin{equation}
\label{eq:hse}
    p_i = p_{i-1} + \frac{1}{2} \Delta z (\rho_i + \rho_{i-1}) \gb \cdot \hat{\bf z}
\end{equation}
and holding the temperature constant in the ghost cells.  This is solved
together with the equation of state.  The velocity is reflected at this boundary.
This procedure follows the form described in \citet{ppm-hse}.  The
left boundary is reflecting and the right boundary is a zero-gradient
outflow.  The top boundary sets the state to simply the conditions
in our outer buffer region of the initial model (see below), with the
normal velocity set to the larger of zero or the velocity at the top
of the domain (this prevents incoming velocities at the top) and the
transverse velocities set to zero.

@@

%======================================================================
% Initial Model
%======================================================================
\section{Initial Model}\label{Sec:inital_model}
\explain{We've updated the equations in this section to use cylindrical
coordinates.}

@@

\added{
\citet{cavecchi:2013} ran simulations at rotation rates of $450$ Hz exploring a range of opacities.
Accounting for the boosting, we would expect to see our 10/10 flame propagate at $\sim 2.25$ times
the speed of these $450$ Hz runs, but instead the speeds range from $4.5$ to $7$ times those reported in
\citet{cavecchi:2013}. The additional boost may come from cyclic fluid motions that would have been
disallowed in the $450$ Hz runs by the enforcement of vertical hydrostatic equilibrium (we already saw
that hydrodynamic effects can accelerate the burning front in our comparison with the 1D flames). The
inclusion of reaction chains beyond triple-$\alpha$ may also increase energy generation and consequently the
propagation speed.}
